---
title: "Barangay Del Rosario Roadside Trees"
author: "ADNU Regis Hub"
date: "9/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### This document is a brief summary of the partial tree inventory along the main roads in Naga City.The data described in this paper came from the fieldwork activities conducted from **December 05, 2017** to **May 23, 2019** and included portions of the 17 barangays in Naga City.  
### Fieldwork activities for this joint project were participated by volunteers from **They Grey We Green, SULOG (Sumaro sa Salog), private individuals, environmental management students, citizen scientists, and headed by the Institute for Environmental Conservation and Research of the Ateneo de Naga University (ADNU-INECAR)**.  
### Tree census was facilitated using a Kobo Collect android application developed in part by the INECAR team. Using the application, the volunteers geotag the trees and input important information from each of the trees like the species or local name and the barangay where it is located. In addition, the volunteers also took pictures of the trees to aid in the validation of species. After fieldwork, the volunteers send their files online to a cloud server that can be accessed through a web application, the KoboToolbox.

### Data from the server was exported and downloaded for data cleaning and processing using R and RStudio. This document was also created using the said software. 

## **Data Summary**

```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'hide', eval=TRUE}
library(dplyr)
library(data.table)
library(reshape2)
library(lubridate)
library(stringr)
```

### As of May 23, 2019, a total of 2082 trees along the roadside were counted in the Project.
#### **Please note that these tree data is PARTIAL as the tree census is far from over and was temporarily discontinued due to the lack of manpower and the onset of the COVID19 Pandemic**.  

### The latest Tree Inventory covered portions of the 17 barangays and the trees counted are listed below. For now, the top three barangays with the most number of trees are **Pacol, San Felipe, and Concepcion Pequena**.  

```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'hide', eval=TRUE}

library(data.table)
test <- fread("test.csv")
barangays <- test[, .N, by = test$barangay]
as.data.table(barangays)
setkey(barangays, N)
barangays2 <- as.data.table(setorder(barangays, -N))
barangays2
```  
```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'hide', eval=TRUE}

barangays2
```  
```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'show', eval=TRUE}

library(plotly)

brgy <- plot_ly(barangays2, y = barangays2$test, x = barangays2$N, type = 'bar', orientation = 'h')
brgy

``` 

## **The Case for Barangay Del Rosario**  

### Most common roadside trees in Barangay Del Rosario are shown by the following graph:  

```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'hide', eval=TRUE}

delrosario<- test[ barangay == "Del Rosario", .N, by = "characteristics"]
as.data.table(setorder(delrosario, -N))

```  

```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'show', eval=TRUE}

library(plotly)
delrosario <- as.data.table(setorder(delrosario, -N))
fig <- plot_ly(delrosario, y = delrosario$characteristics, x = delrosario$N, type = 'bar', orientation = 'h')
fig

```  

### Based on the dataset, currently, the total number of trees that the Census has counted in Barangay Del Rosario is **146**


```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'hide', eval=TRUE}
library(dplyr)
library(data.table)
library(reshape2)
library(lubridate)
library(stringr)
library(leaflet)
```  

```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'hide', eval=TRUE}
delros2 <- test[barangay == "Del Rosario"]
head(delros2)
```  

## Map of Plotted Trees in Barangay Del Rosario:  
### The volunteers gathered data from the roadside trees in the barangay as seen in the map below. Please note that the team used at most, plus or minus three (3) meters as margin of error for the acquisition of tree location (GPS coordinates). On the other hand, one may see that some trees are plotted faraway from the roadside which  maybe due to poor GPS reception or outdated settings in the mobile application used in the early days of the Project.


```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'show', eval=TRUE}

m <- leaflet(data = delros2) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(~longitude, ~latitude, popup= ~as.character(characteristics))
m


```

## How large are the trees along the roadside of Barangay Del Rosario?  
### To visualize this, the map shows the trees and their measured girth at breast height/GBH (circumference of the tree at 1.3 meters) in meters as pop up when the tree bubbles are clicked:

```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'show', eval=TRUE}

delros2 <- delros2 %>% 
  mutate(popup= paste(delros2$characteristics, delros2$gbh, sep = " "))


n <- leaflet(data = delros2) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(~longitude, ~latitude, popup= ~as.character(popup))
n


```  

### It is important to note that the range of tree circumference (at GBH) recorded for the 146 trees is **0.16 to 4.8 ** meters.

```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'show', eval=TRUE}
summary(delros2$gbh)
```    
### Also, it is interesting to note that even though the team were not able to gather the tree height data on all of the trees, the range of height for a small sample is **2.350 to 13.62 ** meters.

```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'show', eval=TRUE}
summary(delros2$height)
```  

## **Tree Benefits (Ecosystem Services) that we freely enjoy here in Naga City**  

### Before we proceed to our brief discussion on the benefits of Trees in an urban area, please take a look at the  summary of the Del Rosario tree dataset:  

```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'show', eval=TRUE}
summary(delros2)
```  
### and the following parameters collected:
```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'show', eval=TRUE}
colnames(delros2)
```  

### The data above is a subset of the whole tree data set of the Trees in the City Project that volunteers and citizen scientists had contributed by doing actual fieldwork under the sun or rain. With this, **the authors are truly grateful to the volunteers and sponsors**.  
## --------------------------------------------------------------------

  
  
## The Benefits of Trees


### Aside from the **beauty and shade these Urban Trees provide**, the trees'  most important ecosystem services are tied to its ability to **photosynthesize**, that is, its ability to absorb carbon dioxide, take in water and nutrients from the soil, make food for itself (carbohydrates and sugars), and give off oxygen and water.
### **Trees are able to lessen flooding and soil erosion** as the trees' roots are continously absorbing water and holding our precious soil.
### **Trees help control and stabilize our microclimate within the City and lessen the impact of the Urban Heat Island Effect**. Simply, **trees are our natural air conditioning systems**. Learn more about this in an actual study conducted in Naga. [Link here](https://www.researchgate.net/publication/336133955_Microclimate_Variability_Analysis_and_a_Proposed_Management_Plan_for_the_Urban_Heat_Reduction_of_Naga_City_Camarines_Sur) 

### Ever wonder why we feel a lot cooler under the shade of the tree or why we usually park our cars under the shade of a tree? Because aside from the leaves and thick foliage that protect us from direct sunlight, the water vapor that the tree produces through photosynthesis, escapes in the small leaf openings called stomata (stomates).  Because of the evaporation of these water vapor, we feel a significantly lower temperature under a tree.   

### **Trees absorb Carbon Dioxide** and store most of them in their bodies as wood and biomass. Because of this, maintaining our existing trees and planting more of them helps in our fight against the effects of the changing climate, that we now call as Climate Emergency.  

### In line with this, **Trees clean the air we breathe** by absorbing or removing pollutants in the air. Do you know that Oxygen, which is the air we breathe, and Water, which is essential in our daily life, are just the "waste materials" of a plant's photosynthesis?  

### On the other hand, Carbon Dioxide , which we humans and animals exhale, is in turn needed by the plants to live and grow. Too bad that in the present, the amount of greenhouse gases including Carbon Dioxide in the atmosphere has far exceeded the ability of plants and plankton to take them up, causing more complicated problems.  

### When we cut the trees and allow them to rot or burn, most of the the stored carbon that they successfully absorbed while they were still alive, go back to the atmosphere, adding more carbon dioxide and greenhouse gases that make our planet hotter and hotter each year.  

### **Trees serve as habitats of other organisms** thus enabling a wildlife to thrive even in cities and urban areas. Most of these trees in our city support a wide variety of wildlife such as orchids and ferns, birds, bees, butterflies, and insects. In nature, an ecosystem with more diverse the species, it is more likely to become more stable and resilient from disturbance.  Like the trees, these organisms also bring us benefits such as pollination and pest control to name a few. Trees promote biodiversity in urban areas. See for example the data on Barangay Del Rosario trees as hosts or habitats of plants and animals:  


```{r, echo = FALSE, warning = FALSE, error=FALSE, message=FALSE, results= 'show', eval=TRUE}
summary(delros2$host)
```  

### **36** out of the inventoried trees serve as host for other wildlife. 

## --------------------------------------------------------------------
## There are  still many more reasons why trees are important inside our cities and urban areas, but we hope that the reasons cited above will help us realize their importance in our daily lives.  
## --------------------------------------------------------------------
  
  
  
  
    
            

### REFERENCES

#### R Core Team (2017). R: A language and environment for statistical computing. R  Foundation for Statistical Computing, Vienna, Austria. URLhttps://www.R-project.org/

#### packages used: dplyr, data.table, lubridate, reshape2, stringr, rmarkdown, plotly, and leaflet*   

#### H. Wickham. ggplot2: Elegant Graphics for Data Analysis.      Springer-Verlag New York, 2016    

#### Hadley Wickham (2019). stringr: Simple, Consistent Wrappers for Common String Operations. R package version 1.4.0. https://CRAN.R-project.org/package=stringr  
   
#### Matt Dowle and Arun Srinivasan (2019). data.table: Extension of data.frame`. R package version 1.12.8. https://CRAN.R-project.org/package=data.table

#### JJ Allaire and Yihui Xie and Jonathan McPherson and JavierLuraschi and Kevin Ushey and Aron Atkins and Hadley Wickham and Joe Cheng and Winston Chang and Richard Iannone (2020).rmarkdown: Dynamic Documents for R. R package version 2.1. URLhttps://rmarkdown.rstudio.com.

#### Hadley Wickham, Romain Fran?ois, Lionel Henry and Kirill M?ller (2020). dplyr: A Grammar of Data Manipulation. R package version 1.0.0.https://CRAN.R-project.org/package=dplyr    

#### Anne Cori (2020). EpiEstim: Estimate Time Varying Reproduction Numbers from Epidemic Curves. R package version 2.2-3. https://CRAN.R-project.org/package=EpiEstim  

#### VP Nagraj, Thibaut Jombart, Nistara Randhawa, Bertrand Sudre, Finlay Campbell and Thomas Crellen (2017). epicontacts: Handling, Visualisation and Analysis of Epidemiological Contacts. R package version 1.1.0. https://CRAN.R-project.org/package=epicontacts    
  
#### Garrett Grolemund, Hadley Wickham (2011). Dates and Times Made Easy with lubridate. Journal of Statistical Software, 40(3), 1-25. URL http://www.jstatsoft.org/v40/i03/    

#### Hadley Wickham (2007). Reshaping Data with the reshape Package. Journal of Statistical Software, 21(12), 1-20. URL http://www.jstatsoft.org/v21/i12/    

#### Joe Cheng, Bhaskar Karambelkar and Yihui Xie (2019). leaflet:Create Interactive Web Maps with the JavaScript 'Leaflet'Library. R package version 2.0.3. https://CRAN.R-project.org/package=leaflet  

#### Yihui Xie (2020). knitr: A General-Purpose Package for Dynamic Report Generation in R. R package version 1.29.    

#### Yihui Xie (2015) Dynamic Documents with R and knitr. 2nd edition. Chapman and Hall/CRC. ISBN 978-1498716963  

#### Yihui Xie (2014) knitr: A Comprehensive Tool for Reproducible Research in R. In Victoria Stodden, Friedrich Leisch and Roger D. Peng, editors, Implementing Reproducible Computational Research. Chapman and Hall/CRC. ISBN 978-1466561595 

